---
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      cache.lazy = FALSE,
                      # force output to LaTeX (which forces the
                      # imposition of fig.pos) and proper figure alignment
                      fig.align = 'center',
                      fig.pos = 'H')

library(dplyr)
library(here)
library(tidyr)
library(stringr)
library(forcats)
library(readr)

library(kableExtra)

library(gridExtra)

library(ggplot2)
extrafont::loadfonts(quiet = TRUE)
library(GGally)
library(hrbrthemes)
theme_set(theme_ipsum())
library(atheylab.utils)

knitr::knit_hooks$set(inline = function(x) {
  atheylab.utils::pretty_print_inline(x, decimals = 2)
})

library(texreg)
```

---
title: "Channel Classification Analysis"
---

Our current method of classifying channels by political leaning is pretty rough: let's take a look at how many we are able to catch.

```{r data_import}
# read in the classification
channel_classification <-  read_csv(here('data/derived_data/analysis_redo/channel_classification.csv'))

# get a list of all the channels we visited
read_csv(here('data/derived_data/analysis_redo/video_info.csv')) %>%
  filter(!is.na(channel)) %>%
  mutate(channel = tolower(channel)) -> video_info

video_info %>%
  count(channel) %>%
  arrange(-n) -> video_channels

# join them together
video_channels %>%
  full_join(channel_classification, by='channel') -> class_full

```

First a summary table: how many R/L/C/missing channels do we have?

```{r summ_tbl}
class_full %>%
  mutate(leaning = ifelse(is.na(leaning), 'unclassified', leaning)) %>%
  count(leaning) %>%
  mutate(share = n / sum(n)) -> summ_tbl

summ_tbl %>%
  mutate(share = scales::percent(share))
```

Anything but a pie chart:

```{r summ_viz}
summ_tbl %>%
  ggplot() + 
  geom_col(aes('', share, fill = leaning)) + 
  scale_fill_manual(values = c("purple", "blue", "red", "gray")) +
  ggtitle('Channel Classification Share')
```

What if we take into account the number of videos we have from each? R/L/C/missing weighted by number of videos from each channel.

```{r}
class_full %>%
  mutate(leaning = ifelse(is.na(leaning), 'unclassified', leaning)) %>%
  group_by(leaning) %>%
  summarise(n = sum(n, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(share = n / sum(n)) -> summ_tbl_alt

summ_tbl_alt %>%
  mutate(share = scales::percent(share))

```

```{r}
summ_tbl_alt %>%
  ggplot() + 
  geom_col(aes('', share, fill = leaning)) + 
  scale_fill_manual(values = c("purple", "blue", "red", "gray")) + 
  ggtitle('Channel Classification Share Weighted by Video Count')

```


